### 2015-09-15 vmcai-2016-tex/bisimulation.tex
Added new define equals macro.

The theorems in this section establish the soundness and
completeness of the symbolic heap approach with respect to Generalized
Symbolic Execution (\gsetxt{}). Intuitively, the theorems imply that any properties proven
with~\gsetxt{} can also be proved using the symbolic heap
algorithm. What follows is a brief description of the requisite terminology,
followed by the theorem statements. The proofs for the theorems, as well as
the complete set of semantic rules for~\gsetxt{} and~\symtxt{} can
be found in ~\cite{Hillery:2015}.

The theorems assert the existence of a
bisimulation between sets of states related by~\gsetxt{} ($p \rgse
p^\prime$) and states related by the symbolic heap and update rules in this paper ($q \rsym
q^\prime$). The relations are on the universe of \emph{well-formed}
states $S$, which have the properties in \secref{sec:sh} with the
constraint that the states are \emph{feasible}: successors exist
unless at \cfgt{end}. Let $p \rgse p^\prime$ be a union over relations
for~\gsetxt{} (see~\cite{Hillery:2015}): $\rgse = \rgse^\mathit{A} \cup
\rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup
\rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup
\rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access
after evaluating the expression for the base reference $\lp
\cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp$, \emph{W} is
a field write after evaluating the expression for the right operand
$\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x^Y$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x^Y P^\prime \Longleftrightarrow \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x^Y p^\prime \Leftrightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose a set $P$ contains a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in~\gsetxt{}: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.

The functional equivalence between heaps in states $p$
and $q$ requires both a mapping to relate the two heaps and a
constraint on the feasibility of that mapping in the presence of a
path constraint from symbolic execution. Subscripts indicate state tuple
members as in $p = (
\cfgnt{L}_p\ \cfgnt{R}_p\ \phi_p\ \eta_p\ \cfgnt{e}_p\ \cfgnt{k}_p )$.
\begin{definition}
\label{def:homomorphism}
A \textbf{homomorphism}, given the universe of field indices $\mathcal{F}$ and the universe of locations $\mathcal{L}$, is 
$$
\begin{array}{l}
 s_p \rightharpoonup_{h} s_q \Leftrightarrow 
\exists h: \mathcal{L} \mapsto \mathcal{L}\ (\forall \cfgnt{l}_\alpha\ (\forall \cfgnt{l}_\beta\ (\forall f \in \mathcal{F}\ ( \forall \phi_\alpha\ (\\ 
\ \ \ \ \ \ \ \ (\phi_\alpha\ \cfgnt{l}_\alpha) \in \cfgnt{L}_p(\cfgnt{R}_p (\cfgnt{l}_\beta,f )) \Rightarrow \\
\ \ \ \ \ \ \ \ \exists \phi_\beta\ ( (\phi_\beta\ h(\cfgnt{l}_\alpha))\in \cfgnt{L}_q(\cfgnt{R}_q (h(\cfgnt{l}_\beta),f ))\ 
 )) ) ) ) )
\end{array}
$$
\end{definition}
\begin{definition}
\label{def:hc}
The \textbf{homomorphism constraint} is
%\begin{align*}
\[
\mathbb{HC}(p \rightharpoonup_{h} q) = 
\bigwedge \{ \phi_b\ | \exists (\phi_a\ l) \in \cfgnt{L}_p^\rightarrow ( (\phi_b\ h(l)) \in \cfgnt{L}_q^\rightarrow)\}
\]
%\end{align*}
\end{definition}
The notation $\cfgnt{L}_p^\rightarrow$ denotes the image of $\cfgnt{L}_p$. Functional equivalence asserts a common structure in the two heaps under certain
conditions. It is used to relate states in $p \rgse
p^\prime$ to states in $q \rsym q^\prime$.
\begin{definition}
\label{representation}
States $(p\ q)$ are in the \textbf{representation relation}, $p \sqsubset q$, if and only if, $\eta_p = \eta_q ,\ \cfgnt{e}_p =
\cfgnt{e}_q ,\ \cfgnt{k}_p = \cfgnt{k}_q$, and there exists a
homomorphism $p \rightharpoonup_{h} q$
such that $\mathbb{S}( \phi_q \wedge \mathbb{HC}(s_p \rightharpoonup_{h} s_q) )$.
The represented relation is extended to sets of states $P$ and a single state $q$ as
$P \sqsubset q \Longleftrightarrow \forall p\ (p \sqsubset q \Leftrightarrow p \in P)$.
\end{definition}
The statement $p \sqsubset q$ ensures that a functionally equivalent
heap to the one in $p$ is present, by the homomorphism, and valid, by
the heap constraint and path constraint, in $q$. As the states in $P$
are only differentiated by heaps and those states are only
differentiated from $q$ by both the heap and path constraint in $q$,
$P \sqsubset q$ implies that $q$ is representative of all the states
in $P$ up to the given point of execution expressed in $\phi_q$.
\begin{definition}
\label{bisimulation}
The \textbf{functional associated to bisimulation} applied to $\sqsubset$, denoted as $F_\sim(\sqsubset)$, is the set of all pairs
$(P\ q)$ such that
\begin{equation}
\label{eqn:BisimulationForwards}
\forall P^\prime\ ( P \rsgse P^\prime \Rightarrow \exists q^\prime\ ( q \rsym q^\prime \wedge P^\prime\ \sqsubset\ q^\prime))
\end{equation}
\begin{equation}
\label{eqn:BisimulationBackwards}
\forall q^\prime\ ( q \rsym q^\prime\Rightarrow \exists P^\prime\ ( P \rsgse P^\prime \wedge P^\prime\ \sqsubset\ q^\prime))
\end{equation}
If $\sqsubset$ is a bisimulation, then the greatest fixed point of $F_\sim(\sqsubset)$ is the bisimilarity relation denoted by $\sim$.
\end{definition}
Other than the use of a meta-relation, $\rsgse$, \defref{bisimulation} reasons over the typical forward and backward
simulation~\cite{Sangiorgi:2011}.

>>>>

The theorems in this section establish the soundness and
completeness of the symbolic heap approach with respect to Generalized
Symbolic Execution (\gsetxt{}). Intuitively, the theorems imply that any properties proven
with~\gsetxt{} can also be proved using the~\symtxt{}
algorithm. What follows is a brief description of the requisite terminology,
followed by the theorem statements. The proofs for the theorems, as well as
the complete set of semantic rules for~\gsetxt{} and~\symtxt{} can
be found in ~\cite{Hillery:2015}.

The theorems assert the existence of a
bisimulation between sets of states related by~\gsetxt{} ($p \rgse
p^\prime$) and states related by the symbolic heap and update rules in this paper ($q \rsym
q^\prime$). The relations are on the universe of \emph{well-formed}
states $S$, which have the properties in \secref{sec:symheap} with the
constraint that the states are \emph{feasible}: successors exist
unless at \cfgt{end}. Let $p \rgse p^\prime$ be a union over relations
for~\gsetxt{} (see~\cite{Hillery:2015}): $\rgse = \rgse^\mathit{A} \cup
\rgse^\mathit{A^\prime} \cup \rgse^\mathit{W} \cup
\rgse^\mathit{W^\prime} \cup \rgse^\mathit{E} \cup
\rgse^\mathit{E^\prime} \cup \rcom$, where \emph{A} is a field access
after evaluating the expression for the base reference $\lp
\cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp$, \emph{W} is
a field write after evaluating the expression for the right operand
$\lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgnt{*} \rightarrow
\cfgnt{k}\rp$, \emph{E} is a reference compare after evaluating the
left and right operands $\lp \cfgnt{v}\ \cfgt{=}\ \cfgnt{*}
\rightarrow \cfgnt{k}\rp$, the prime symbol indicates a null reference
in the operation or a false outcome, and \emph{\com} is everything
else in the language. Any state relation, say $\rightarrow_x^Y$, is
trivially extended to sets of states as
$$
P \hookrightarrow_x^Y P^\prime \defeq \forall p \in P\ (\forall p^\prime\ (p \rightarrow_x^Y p^\prime \Leftrightarrow p^\prime \in P^\prime))
$$
Let $\hookrightarrow_\gse$ be the extension of $\rgse$ to sets of
states. From this extension, a new meta transition relation is defined
over sets of states as
$$
\rsgse = \hookrightarrow_\gse^\mathit{A}
\cup \hookrightarrow_\gse^\mathit{A^\prime} \cup \hookrightarrow_\gse^\mathit{W} \cup
\hookrightarrow_\gse^\mathit{W^\prime} \cup \hookrightarrow_\gse^\mathit{E} \cup \hookrightarrow_\gse^\mathit{E^\prime}
\cup \hookrightarrow_\com
$$
The relation captures the notion of splitting groups of heaps at
certain operations. For example, suppose a set $P$ contains a single state $p$
with all references uninitialized. If $p$ is a field access state, it has two
potential successors in~\gsetxt{}: a non-null reference and a null reference. 
Thus, the $\rsgse$ relation has two successors and divides $P$ into the two outcomes.

The functional equivalence between heaps in states $p$
and $q$ requires both a mapping to relate the two heaps and a
constraint on the feasibility of that mapping in the presence of a
path constraint from symbolic execution. Subscripts indicate state tuple
members as in $p = (
\cfgnt{L}_p\ \cfgnt{R}_p\ \phi_p\ \eta_p\ \cfgnt{e}_p\ \cfgnt{k}_p )$.
\begin{definition}
\label{def:homomorphism}
A \textbf{homomorphism}, given the universe of field indices $\mathcal{F}$ and the universe of locations $\mathcal{L}$, is 
$$
\begin{array}{l}
 s_p \rightharpoonup_{h} s_q \defeq
\exists h: \mathcal{L} \mapsto \mathcal{L}\ (\forall \cfgnt{l}_\alpha\ (\forall \cfgnt{l}_\beta\ (\forall f \in \mathcal{F}\ ( \forall \phi_\alpha\ (\\ 
\ \ \ \ \ \ \ \ (\phi_\alpha\ \cfgnt{l}_\alpha) \in \cfgnt{L}_p(\cfgnt{R}_p (\cfgnt{l}_\beta,f )) \Rightarrow \\
\ \ \ \ \ \ \ \ \exists \phi_\beta\ ( (\phi_\beta\ h(\cfgnt{l}_\alpha))\in \cfgnt{L}_q(\cfgnt{R}_q (h(\cfgnt{l}_\beta),f ))\ 
 )) ) ) ) )
\end{array}
$$
\end{definition}
\begin{definition}
\label{def:hc}
The \textbf{homomorphism constraint} is
%\begin{align*}
\[
\mathbb{HC}(p \rightharpoonup_{h} q) \defeq
\bigwedge \{ \phi_b\ | \exists (\phi_a\ l) \in \cfgnt{L}_p^\rightarrow ( (\phi_b\ h(l)) \in \cfgnt{L}_q^\rightarrow)\}
\]
%\end{align*}
\end{definition}
The notation $\cfgnt{L}_p^\rightarrow$ denotes the image of $\cfgnt{L}_p$. Functional equivalence asserts a common structure in the two heaps under certain
conditions. It is used to relate states in $p \rgse
p^\prime$ to states in $q \rsym q^\prime$.
\begin{definition}
\label{representation}
States $(p\ q)$ are in the \textbf{representation relation}, $p \sqsubset q$, if and only if, $\eta_p = \eta_q ,\ \cfgnt{e}_p =
\cfgnt{e}_q ,\ \cfgnt{k}_p = \cfgnt{k}_q$, and there exists a
homomorphism $p \rightharpoonup_{h} q$
such that $\mathbb{S}( \phi_q \wedge \mathbb{HC}(s_p \rightharpoonup_{h} s_q) )$.
The represented relation is extended to sets of states $P$ and a single state $q$ as
$P \sqsubset q \defeq \forall p\ (p \sqsubset q \Leftrightarrow p \in P)$.
\end{definition}
The statement $p \sqsubset q$ ensures that a functionally equivalent
heap to the one in $p$ is present, by the homomorphism, and valid, by
the heap constraint and path constraint, in $q$. As the states in $P$
are only differentiated by heaps and those states are only
differentiated from $q$ by both the heap and path constraint in $q$,
$P \sqsubset q$ implies that $q$ is representative of all the states
in $P$ up to the given point of execution expressed in $\phi_q$.
\begin{definition}
\label{bisimulation}
The \textbf{functional associated to bisimulation} applied to $\sqsubset$, denoted as $F_\sim(\sqsubset)$, is the set of all pairs
$(P\ q)$ such that
\begin{equation}
\label{eqn:BisimulationForwards}
\forall P^\prime\ ( P \rsgse P^\prime \Rightarrow \exists q^\prime\ ( q \rsym q^\prime \wedge P^\prime\ \sqsubset\ q^\prime))
\end{equation}
\begin{equation}
\label{eqn:BisimulationBackwards}
\forall q^\prime\ ( q \rsym q^\prime\Rightarrow \exists P^\prime\ ( P \rsgse P^\prime \wedge P^\prime\ \sqsubset\ q^\prime))
\end{equation}
If $\sqsubset$ is a bisimulation, then the greatest fixed point of $F_\sim(\sqsubset)$ is the bisimilarity relation denoted by $\sim$.
\end{definition}
Other than the use of a meta-relation, $\rsgse$, \defref{bisimulation} reasons over the typical forward and backward
simulation~\cite{Sangiorgi:2011}.
